### 2023-12-11 R/run_first/1.pattern_examples_MASCARA_pipeline.Rmd

---
title: "MASCARA Playground"
author: "Fred White"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(devtools)
library(ggplot2)
library(MetStaT)
library(gASCA)
library(patchwork)
library(readxl)
library(ggfortify)
library(ggrepel)
library(grid)
library(gridExtra)
library(tidyr)
library(tidyverse)
library(reshape2)
library(scales)
library(matrixStats)
library(DESeq2)
library(MASS)
library(glmnet)
library(doParallel)    
library(MUVR) 
library(pheatmap)
library(caret)
library(data.table)
library(plotly)
library(mixOmics)
library(DiscriMiner)
library(pracma)
library(matlib)
library(truncnorm)

source("../DATA_SIM_FUNCS.R")

```

```{r}

meta <- cbind.data.frame(rep(c(1,-1), each = 12), rep(c(1:4), each = 3), rep(c(-1:-4,1:4), each = 3))
colnames(meta) <- c("growth_condition","time","interaction")
meta <- as.data.frame(apply(meta,2,factor))  #9

Experiment_responders <- 30

```

```{r}

sim_data <- Create_Core_DEV_2(nreps = 3, meta = meta, 
                        EffectSize = c(X_a_ab = 5, time = 1, E = 10, mu = 1, Struc_E = 1),    # E = 1, Struc_E = 0.4
                        struc_resid = TRUE, e_sigma = c(1.5,0.8,0.0), a_sigma = c(1,0.7,0.6,0.1),
                        b_sigma = c(1,0.9), irr_spikes = TRUE, SCORE_SEED = 12, Experiment_responders = Experiment_responders,
                        struc_seed = 127, score_plot = FALSE, plot = FALSE)

```

```{r}
baits <- paste0("X_",c(1985:1988))   #501:510
spikes <- paste0("X_",c(1989:2000))

ar <- ASCA(sim_data[[1]], m = meta, Baits = baits, spikes = spikes, ncands = ncands, distance_calc = FALSE, Return_Model = FALSE)

```

```{r}

gc <- scale(ar[[4]]$decomposition$growth_condition[,1985:1988])  #0.1*
time <- sim_data$Effects[[2]][,1985:1988] #15 *

int <- scale(sim_data$Effects[[1]][,1985:1988])  #0.08*

ambient <- sim_data$Effects[[4]][,1985:1988]  #1.2 * 
ambient[13:24,] <- 0

noise <- sim_data$Effects[[5]][,1985:1988]  #1.2 *

rownames(gc) <- paste0("X_",1:nrow(gc))
rownames(time) <- paste0("X_",1:nrow(time))
rownames(int) <- paste0("X_",1:nrow(int))
rownames(ambient) <- paste0("X_",1:nrow(ambient))
rownames(noise) <- paste0("X_",1:nrow(noise))

```

```{r}
max <- max(c(gc,time,int,ambient,noise))
min <- min(c(gc,time,int,ambient,noise))

mData <- reshape2::melt(gc)
colnames(mData)[1:2] <- c("Samples","Features")
# mData$Features <- fct_rev(mData$Features)

library(viridis)
library(hrbrthemes)
library(ggplot2)
p1 <- ggplot(mData, aes(Features, Samples, fill= value), show.legend = FALSE) + 
  geom_tile() +
  ggtitle("Growth Condition") +
  scale_fill_viridis(discrete=FALSE) +  
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

mData <- reshape2::melt(time)
colnames(mData)[1:2] <- c("Samples","Features")

library(viridis)
library(hrbrthemes)
library(ggplot2)
p2 <- ggplot(mData, aes(Features, Samples, fill= value)) + 
  geom_tile() +
  ggtitle('Time') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

interaction <- int - time

mData <- reshape2::melt(interaction)
colnames(mData)[1:2] <- c("Samples","Features")

library(viridis)
library(hrbrthemes)
library(ggplot2)
p3 <- ggplot(mData, aes(Features, Samples, fill= value)) + 
  geom_tile() +
  ggtitle('Interaction') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

mData <- reshape2::melt(ambient)
colnames(mData)[1:2] <- c("Samples","Features")

library(viridis)
library(hrbrthemes)
library(ggplot2)
p4 <- ggplot(mData, aes(Features, Samples, fill= value)) + 
  geom_tile() +
  ggtitle('Ambient') +
  scale_fill_viridis(discrete=FALSE ) +   #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

mData <- reshape2::melt(noise)
colnames(mData)[1:2] <- c("Samples","Features")

library(viridis)
library(hrbrthemes)
library(ggplot2)
p5 <- ggplot(mData, aes(Features, Samples, fill= value)) + 
  geom_tile() +
  ggtitle('Residuals') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

data <- (0.1 * interaction) + (1.5*time) + ambient + noise
mData <- reshape2::melt(data)
colnames(mData)[1:2] <- c("Samples","Genes")

library(viridis)
library(hrbrthemes)
library(ggplot2)
p0 <- ggplot(mData, aes(Genes, Samples, fill= value)) + 
  geom_tile() +
  ggtitle('Data') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))

plots <- list(p1, p2, p3, p4, p5)

p5

```

```{r}
ps <- patchwork::patchworkGrob(wrap_plots(plots) )    #+ plot_layout(guides = "collect") -

ps1 <- grid.arrange(ps)

```

```{r}

mData <- reshape2::melt(noise + ambient)
colnames(mData)[1:2] <- c("Samples","Features")
# mData$Features <- fct_rev(mData$Features)

library(viridis)
library(hrbrthemes)
library(ggplot2)
p5 <- ggplot(mData, aes(Features, Samples, fill= value)) + 
  geom_tile() +
  ggtitle('Residuals') +
  scale_fill_viridis(discrete=FALSE) +  #, limits = c(min,max)
  theme_bw() +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

p5
```

```{r}

# 
input <- data
#

```

```{r}

spls_res <- simpls.fit(data.matrix(input[,1:2]), data.matrix(input[,-c(1:2)]), ncomp = 2)

scores <- spls_res$scores

X_P <- spls_res$loadings
Y_P <- spls_res$Yloadings

X_TP <- (spls_res$scores %*% t(spls_res$loadings))
Y_TP <- (spls_res$scores %*% t(spls_res$Yloadings))

X_E <- spls_res$residuals[,1,]
Y_E <- spls_res$residuals[,2,]

```

```{r}
source("../MASCARA_FUNCS.R")

#classic; q bar target projection
MASCARA_2 <- MASCARA4_test_2(ar[[3]], baits, spikes = spikes)
which(rownames(MASCARA_2[[2]]) %in% spikes)

#u1 of baits as pls1 response
MASCARA_3 <- MASCARA4_test_3(ar[[3]], baits, spikes = spikes)
which(rownames(MASCARA_3[[2]]) %in% spikes)

#sr
MASCARA_4 <- MASCARA4_test_7(ar[[3]], baits, spikes = spikes)
which(rownames(MASCARA_4[[2]]) %in% spikes)

#vip
MASCARA_5 <- MASCARA4_test_8(ar[[3]], baits, spikes = spikes)
which(rownames(MASCARA_5[[2]]) %in% spikes)

```

```{r}
#plotting qbar TP for demo

loadings <- MASCARA_2[[4]]$loadings[]
yloadings <- MASCARA_2[[4]]$Yloadings[]

scores <- MASCARA_2[[4]]$scores[]
colnames(scores) <- gsub(" ","_", colnames(scores))

colnames(loadings) <- gsub(" ","_", colnames(loadings))

line <- colMeans(yloadings)

#rotation of scores and loadings for demo purposes only

  #adapted from https://stackoverflow.com/questions/1897704/angle-between-two-vectors-in-r
  angle <- function(x,y){

    x <- c(as.matrix(x))
    y <- c(as.matrix(y))

    dot.prod <- x%*%y 
    norm.x <- norm(x,type="2")
    norm.y <- norm(y,type="2")
    theta <- acos(dot.prod / (norm.x * norm.y))
    return(as.numeric(theta))
  }

  #rotated loadings
  ang <- angle(line,c(1,-0.8))

  rot.mat2 <- matrix(c(cos(ang),-sin(ang),sin(ang),cos(ang)),nrow = 2, byrow = TRUE) #clockwise rotation

  rotated_point2 <- as.data.frame(c(as.matrix(line)) %*% rot.mat2)

  rotate <- function(x,y){
    x <- c(as.matrix(x))
    x %*% y
  }

  loadings_r_baits <- t(apply(loadings,1,rotate, y = rot.mat2))

  scores_r_baits <- t(apply(scores,1,rotate, y = rot.mat2))

  colnames(loadings_r_baits) <- c("Comp_1","Comp_2")
  colnames(scores_r_baits) <- c("Comp_1","Comp_2")

scores_r_baits <- cbind.data.frame(scores_r_baits, meta)

yloadings <- t(apply(yloadings,1,rotate, y = rot.mat2))

```

```{r}
line <- colMeans(yloadings)

LINE <- rbind(line, c(0,0))
LINE <- t(LINE)
colnames(LINE) <- c("to","from")

scores <- cbind.data.frame(scores,meta[,1:2])

a <- ggplot(scores_r_baits, aes(x = Comp_1, y = Comp_2)) +
  geom_point(aes(colour = growth_condition, shape = time, size = 20)) +
  ggtitle("Scores") + 
  theme_bw()

colnames(yloadings) <- colnames(loadings_r_baits)

loadings_r_baits <- rbind(loadings_r_baits, yloadings)
loadings_r_baits <- cbind.data.frame("ID" = rownames(loadings_r_baits), loadings_r_baits)
```

```{r}
loadings_r_baits$ID[-which(loadings_r_baits$ID %in% c(baits,spikes))] <- "Other"

loadings_r_baits$ID[which(loadings_r_baits$ID %in% spikes)] <- "Spikes"
loadings_r_baits$ID[which(loadings_r_baits$ID %in% baits)] <- "Baits"
```

```{r}
b <- ggplot(loadings_r_baits, aes(x = Comp_1, y = Comp_2, colour = ID, shape = ID)) +
  geom_point() +
    ggtitle("Loadings") +

  geom_segment(x = 0, y = 0, xend = line[1], yend = line[2], lineend = "round", linewidth = 2, colour = "purple", arrow = arrow()) +    
  scale_shape_manual(values=c(3, 16, 17))+

    scale_color_manual(values=c('#E69F00',"black", '#56B4E9'))+
      geom_point(data = loadings_r_baits[which(loadings_r_baits$ID %in% c("Baits","Spikes")),], aes(size = 15)) +
  theme_bw()

b
```