### 2024-01-22 R/MASCARA_Tutorial.Rmd
24_01_22_1

---
title: "MASCARA: Coexpression Analysis Tutorial"
author: "Fred White"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

```{r, echo=FALSE}
library(devtools)
library(ggplot2)
library(MetStaT)
library(gASCA)
library(patchwork)
library(readxl)
library(ggfortify)
library(ggrepel)
library(grid)
library(gridExtra)
library(tidyr)
library(tidyverse)
library(reshape2)
library(scales)
library(matrixStats)
library(DESeq2)
library(MASS)
library(glmnet)
library(doParallel)    
library(MUVR) 
library(pheatmap)
library(caret)
library(data.table)
library(plotly)
library(mixOmics)
library(DiscriMiner)
library(pracma)
library(matlib)
library(truncnorm)

source("DATA_SIM_FUNCS.R")
source("MASCARA_FUNCS.R")
```

Here we show a typical use case of MASCARA applied to a simulated dataset.

User must supply a dataframe of filtered + transformed data plus associated metadata.

```{r}

meta <- cbind.data.frame(rep(c(1,-1), each = 12), rep(c(1:4), each = 3), rep(c(-1:-4,1:4), each = 3))
colnames(meta) <- c("growth_condition","time","interaction")
meta <- as.data.frame(apply(meta,2,factor))  #9

Experiment_responders <- 30

set.seed(3)
# rnd_seed <- abs(rnorm(3, mean = 0.7, sd = 0.5))

baits <- paste0("X_",c(1985:1988))   #501:510
spikes <- paste0("X_",c(1989:2000))

sim_data <- Create_Core_DEV_2(nreps = 3, meta = meta, plot = TRUE, 
                        EffectSize = c(X_a_ab = 1.8, time = 1, E = 10, mu = 1, Struc_E = 1), 
                        struc_resid = TRUE, e_sigma = c(1.5,0.8,0.0), a_sigma = c(1,0.7,0.6,0.1),
                        b_sigma = c(1,0.9), irr_spikes = TRUE, SCORE_SEED = 12, 
                        Experiment_responders = Experiment_responders, struc_seed = 127)

```

```{r}
ar <- ASCA(sim_data[[1]], m = meta, Baits = baits, spikes = spikes, ncands = ncands, distance_calc = FALSE, Return_Model = FALSE)
```

```{r}
#classic
MASCARA_2 <- MASCARA4_test_2(ar[[3]], baits, spikes = spikes)
which(rownames(MASCARA_2[[2]]) %in% spikes)
```