### 2025-01-20 R/Caldana_application.Rmd

---
title: "Caldana_metabolomics_application"
author: "Fred White"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
# library(readxl)
library(openxlsx)
library(ggplot2)
library(patchwork)
library(gASCA)
library(pls)
library(pracma)
library(pheatmap)

source("DATA_SIM_FUNCS.R")
source("MASCARA_FUNCS.R")
```

```{r}
data <- as.data.frame(read.xlsx("Data_Caldana/Caldana-et-al_Normalized-metabolic-data.xlsx"))

```

```{r}
# meta <- cbind.data.frame(colnames(data), t(data[1:2,]))
# meta <- meta[-1,]

# meta$Temperature <- gsub("-.*","",meta$Temperature)
# 
# meta$Light <- gsub(".*-","",meta$Light)

meta <- read.table("Caldana_meta.csv", sep = ",", header = T)
meta$ID <- paste(meta$Temperature, meta$Light, meta$Time, meta$Replicate, sep = "_")
```

```{r, warning=FALSE}
data <- data[-c(1,2),]
Metabolites <- data[,1]
data <- (apply(data[,-c(1)],2, as.numeric))
rownames(data) <- Metabolites

colnames(data) <- meta$ID

data[which(is.na(data), arr.ind = T)] <- 0
```

```{r}
#filtering?

```

```{r}
data <- log2(data + 1)

# data <- data[-grep("Similar.to.Adenine..2TMS",rownames(data)),]
# Metabolites <- Metabolites[-grep("Similar.to.Adenine..2TMS",Metabolites)]
# data <- data[-grep("Similar.to.Adenine..2TMS",rownames(data)),]
# Metabolites <- Metabolites[-grep("Similar.to.Adenine..2TMS",Metabolites)]
```

```{r}
PCA <- prcomp(t(data))#, scale = F, center = F)

A <- PCA$x
B <- PCA$rotation

df <- cbind.data.frame(meta[,c("Temperature","Light","Time")], A)

df$Light <- factor(df$Light, levels = c("D","LL","L","HL"))
# colnames(df)[-which(colnames(df) %in% colnames(meta))] <- paste0("PC",colnames(df)[-which(colnames(df) %in% colnames(meta))])

# pdf(paste0(Sys.Date(),"_feature_loadings_project_full2.pdf"))
scores <- ggplot(df, aes(x = PC1,y = PC2, colour = Temperature, shape = (Light), size = Time)) +
  geom_point(alpha = 0.5) +
  ggtitle("Scores") +
  theme_bw()

#CHANGE

df <- cbind.data.frame(Metabolites,B)

# do.call(c,lapply(Data,colnames))

# pdf(paste0(Sys.Date(),"_feature_loadings_project_full2.pdf"))
loadings <- ggplot(df, aes(x = PC1,y = PC2, alpha = 0.5)) +
  geom_point(alpha = 0.5) + 
  scale_color_manual(values = colours) +
  ggtitle("Loadings") +
  theme_bw()

scores + loadings  + plot_annotation(title = "PCA")

```

```{r}
# #PERMANOVA
# options(knitr.kable.NA = '')
# library(vegan)
# #calculate euclidean distance between samples
# #this matrix should be of the dimensions: number of samples * number of samples
# 
# dist_matrix <- dist(t(data))
# 
# 
# 
# permanova_re <- adonis2(dist_matrix ~ Temperature * Light * Time, 
#                             meta[,c("Temperature","Light","Time")], permutations = 999, method = "euclidean", 
#                             sqrt.dist = FALSE, add = FALSE, by = "terms", 
#                             parallel = 4)
# 
# 
# 
# kable(permanova_re)
#

```

We calculate an ASCA model on the data to partition the experimental and residual variance:

```{r}
ar <- ASCA_decompose(d = apply(meta[,-c(4,5)],2,as.factor), x = t(data), f = "Temperature * Light * Time")

#unexpectedly different from PERMANOVA R2
norm(ar$residuals, type = "F")

norm(ar$decomposition$Temperature, type = "F")

resids <- ar$residuals
```

```{r}
#select some baits (based on PESCA model ?..)

# cors <-

p1 <- pheatmap(cov(resids), show_colnames = F, fontsize_row = 8)
```

```{r}

p2 <- pheatmap(cov(t(data[p1$tree_row$order,])), show_colnames = F, fontsize_row = 8, cluster_rows = F, cluster_cols = F)

# p2 <- pheatmap(cor(t(data[])), show_colnames = F, fontsize_row = 8)

```

Observe the clustering of certain amino acids with high residual correlations, for example Lysine, Tyrosine, Leucine and Isoleucine.

```{r}
baits <- c("Lysine","Tyrosine","Leucine")

```

```{r}
Y_hat <-  t(data) - resids

# time plays a larger role in caldana data -
# Y_hat_nt <-  t(data) - resids - ar$decomposition$Time

Y_hat_r <- cov(t(data))

E_r <- cov(resids)

# plot(Y_hat_r[upper.tri(Y_hat_r)], E_r[upper.tri(E_r)], xlab = "Between Correlations", ylab = "Within Correlations")
# 
# plot(Y_hat_r[], E_r[], xlab = "Between Correlations", ylab = "Within Correlations")

#alpha by norm(resid)/norm(data-resid)

#shape by quantiles below

ratio <- resids^2/Y_hat^2

#make this signless ?
ratio2 <- abs(E_r)/abs(Y_hat_r)

```

```{r}
library(reshape2)

df <- cbind.data.frame(melt(Y_hat_r), melt(E_r)[,3])

colnames(df)[c(3:4)] <- c("Total","Within")

df$Ratio <- abs(df$Within)/abs(df$Total)

df$Ratio2 <- abs(df$Total)/abs(df$Within)

df <- df[-which(df$Var1 == df$Var2),]

#subset higher within examples
df$Covariance <- df$Within

#subset higher Total examples
df$Covariance[which(abs(df$Total) > quantile(abs(df$Total),0.99))] <- "Total"

df$Covariance[which(abs(df$Within) > quantile(abs(df$Within),0.99))] <- "Within"

df$Covariance[-which(df$Covariance %in% c("Within","Total"))] <- "none"

```

```{r}
ggplot(df, aes(x = Total, y = Within)) +
  geom_point(aes(alpha = 0.1)) +
  geom_point(data = df[which(df$Covariance %in% c("Within","Total")),],
aes(alpha = 1, size = 1.2, colour = Covariance)) +
  scale_color_manual(values = c("lightgreen","purple")) + 
     # geom_abline(intercept = 0,slope = 1)
  guides(alpha = "none", size = "none") +
  theme_bw()

```

```{r}
head(df[order(abs(df$Within), decreasing = T),],20)
head(df[order(abs(df$Total), decreasing = T),],20)
df[which(df$Covariance == "Total"),]

```

```{r}
which(ratio2 > quantile(ratio2,0.999),arr.ind = T)
which(ratio2 < quantile(ratio2,0.001),arr.ind = T)

#check sign or signless ratio

#make above plot but make the high ratio standout
```

```{r}
#MASCARA PLOTS

```

Before continuing it is a good idea to check that the set of baits are indeed correlated in the residual variance:

```{r}
cor(resids[,which(colnames(resids) %in% baits)])
```
In this case they are, in the case of real data we expect these values to be somewhat lower (values ~ 0.5 are reason enough to continue).

Another sanity check at this stage is to look at the baits residual variance against each other:

```{r}

plot(resids[,baits[1]], resids[,baits[2]])

```

We can see above that there is a relationship between the examples.

From here we take the residuals and calculate a PLS2 model with a predefined set of baits using the SIMPLS algorithm:

```{r}

spls_res <- simpls.fit(resids[,-which(colnames(resids) %in% baits)],
                          resids[,which(colnames(resids) %in% baits)], ncomp = 2)

```

Selecting features with $\mathbf{\bar{q}}$ target projection:

```{r}
q_bar <- colMeans(spls_res$Yloadings)
R <- spls_res$projection
R_TP <- (dot(t(R),q_bar)/dot(q_bar,q_bar))

Candidates <- as.data.frame(R_TP[order(R_TP, decreasing = TRUE)])
```

Now we have our ordered candidate list, top of the list represents features that have the strongest positive association with the baits.