### 2025-01-20 R/MASCARA_Tutorial.Rmd

---
title: "MASCARA: Coexpression Analysis Tutorial"
author: "Fred White"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

```{r, echo=FALSE}
library(devtools)
library(ggplot2)
library(MetStaT)
library(gASCA)
library(patchwork)
library(readxl)
library(ggfortify)
library(ggrepel)
library(grid)
library(gridExtra)
library(tidyr)
library(tidyverse)
library(reshape2)
library(scales)
library(matrixStats)
library(DESeq2)
library(MASS)
library(glmnet)
library(doParallel)    
# library(MUVR) 
library(pheatmap)
library(caret)
library(data.table)
library(plotly)
# library(mixOmics)
# library(DiscriMiner)
library(pracma)
library(matlib)
library(truncnorm)

source("DATA_SIM_FUNCS.R")
source("MASCARA_FUNCS.R")
```

Here we show a typical use case of MASCARA applied to a simulated dataset.

User must supply a dataframe of filtered + transformed data plus associated metadata, with samples in the rows and genes in the columns.

```{r}
meta <- cbind.data.frame(rep(c(1,-1), each = 12), rep(c(1:4), each = 3), rep(c(-1:-4,1:4), each = 3))
colnames(meta) <- c("growth_condition","time","interaction")
meta <- as.data.frame(apply(meta,2,factor))  
knitr::kable(head(meta))
```

Data is assumed to be approximately normally distributed:

```{r setting simulation parameters}
Experiment_responders <- 30

set.seed(3)

baits <- paste0("X_",c(1985:1988))  
spikes <- paste0("X_",c(1989:2000))

sim_data <- Create_Core_DEV_2(nreps = 3, meta = meta, plot = TRUE, 
                        EffectSize = c(X_a_ab = 1.8, time = 1, E = 10, mu = 1, Struc_E = 1), 
                        struc_resid = TRUE, e_sigma = c(1.5,0.8,0.0), a_sigma = c(1,0.7,0.6,0.1),
                        b_sigma = c(1,0.9), irr_spikes = TRUE, SCORE_SEED = 12, 
                        Experiment_responders = Experiment_responders, struc_seed = 127)[[1]]
hist(sim_data)
```

We calculate an ASCA model on the data to partition the experimental and residual variance:

```{r}
ar <- ASCA_decompose(d = meta[,1:2], x = sim_data, f = "growth_condition + time + growth_condition:time")
resids <- ar$residuals
```

Before continuing it is a good idea to check that the set of baits are indeed correlated in the residual variance:

```{r}
cor(resids[,which(colnames(resids) %in% baits)])
```
In this case they are, in the case of real data we expect these values to be somewhat lower (values ~ 0.6 are reason enough to continue).

Another sanity check at this stage is to look at the baits residual variance against eachother:

```{r}

plot(resids[,baits[1]], resids[,baits[2]])

```

We can see above that there is a relationship between the examples.

From here we take the residuals and calculate a PLS2 model with a predefined set of baits using the SIMPLS algorithm:

```{r}

spls_res <- simpls.fit(resids[,-which(colnames(resids) %in% baits)],
                          resids[,which(colnames(resids) %in% baits)], ncomp = 2)

```

Selecting features with $\mathbf{\bar{q}}$ target projection:

```{r}
q_bar <- colMeans(spls_res$Yloadings)
R <- spls_res$projection
R_TP <- (dot(t(R),q_bar)/dot(q_bar,q_bar))

Candidates <- as.data.frame(R_TP[order(R_TP, decreasing = TRUE)])
```

Now we have our ordered candidate list, top of the list represents features that have the strongest positive association with the baits.