### 2025-01-20 R/MASCARA_b_s_sim.Rmd

---
title: "sim_b_s"
author: "Fred White"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r , echo=FALSE, include=FALSE}
source("DATA_SIM_FUNCS.R")
source("MASCARA_FUNCS.R")
```

```{r, echo = FALSE, include=FALSE}
library(devtools)
library(ggplot2)
# library(MetStaT)
library(gASCA)
library(patchwork)
library(readxl)
library(ggfortify)
library(ggrepel)
library(grid)
library(gridExtra)
library(tidyr)
library(tidyverse)
library(reshape2)
library(scales)
library(matrixStats)
library(DESeq2)
library(MASS)
library(glmnet)
library(doParallel)
# library(MUVR)
library(pheatmap)
# library(caret)
# library(mixOmics)
library(cowplot)
library(reticulate)
library(biomformat)
# library(DiscriMiner)
library(viridis)
library(directlabels)
library(ggpmisc)
library(DescTools)
library(data.table)
library(plotly)
# library(ggbreak)
# library(rhmmer)
library(pracma)
library(pls)
```

```{r,echo = F}

# Noise_Tests <- readRDS("Noise_Sim_Coexp_24_01_22.RDS")
# getwd()
Noise_Tests <- readRDS("Noise_Sim_b_s_24_12_04.RDS")    #Noise_Sim_Coexp_24_05_24.RDS
# Noise_Tests <- Noise_Tests[1:4]
names(Noise_Tests) <- c("ASCA", "PLS", "Correlations", "MASCARA")

Noise_Tests <- Noise_Tests[c("ASCA", "PLS", "Correlations", "MASCARA")]
Noise_Tests <- lapply(Noise_Tests, narm, keep_col1 = T)

bx <- c(2:15)
sx <-c(2:50)

POI <- expand_grid(bx,sx)

s_B_res <- data.matrix(do.call(rbind,lapply(Noise_Tests, function(x) x[1,])))

s_B_res <- cbind.data.frame(POI, t(s_B_res))
s_B_res <- melt(s_B_res, id.vars = c("bx","sx"))
colnames(s_B_res)[3:4] <- c("Method","GMR")

  i <- NULL
for(i in 1:nrow(s_B_res)){

  s_B_res$GMR[i] <- s_B_res$GMR[i]/ (prod(c(1:s_B_res$sx[i]))^(1/s_B_res$sx[i]))

}

s_B_res$GMR <- log2(s_B_res$GMR)
max_value <- max(s_B_res$GMR)
min_value <- min(s_B_res$GMR)

library(viridis)

ggplot(s_B_res, aes(x = bx, y = sx)) +
  geom_tile(aes(fill = GMR)) +
  # scale_fill_gradient(low = "blue", high = "yellow", limits = c(min_value,max_value))+
  scale_fill_viridis_c(limits = c(min_value, max_value)) +
  facet_wrap(~ Method) +
  labs(title = "Num baits & spikes",
       x = "Baits",
       y = "Spikes",
       color = "GMR")

```