### 2025-01-20 R/WGCNA_testing.Rmd

---
title: "WGCNA_test"
author: "Fred White"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(WGCNA)
library(matrixStats)
library(gASCA)
library(pracma)
source("DATA_SIM_FUNCS.R")
#getwd()
```

```{r}
Create_Core_DEV <- Create_Core_DEV_2

baits <- paste0("X_",c(1985:1988))   #501:510
spikes <- paste0("X_",c(1989:2000))

meta <- cbind.data.frame(rep(c(1,-1), each = 12), rep(c(1:4), each = 3), rep(c(-1:-4,1:4), each = 3))
colnames(meta) <- c("growth_condition","time","interaction")

```

```{r}

y <- rep(c(0.0004,0.002, 0.01),3)

x <- rep(30,9)  #100

nse <- rep(c(0.1,0.3,0.5), each = 3)

# y <- c(0.1, 2, 4)
# x <- c(25,125,250)

main_lev <- x*y
ERs <- x

nreps <- 3

a <- 6
i <- 19
    sim_data <- Create_Core_DEV(nreps = nreps, meta = meta, plot = FALSE, 
                    EffectSize = c(X_a_ab = main_lev[a], time = 0.2, E = nse[a], mu = 0.2, Struc_E = 0.05), 
                    struc_resid = T, e_sigma = c(1.5,0.8,0.0), a_sigma = c(1.5,0.7,0.6,0.1),
                    b_sigma = c(1,0), irr_spikes = FALSE, SCORE_SEED = (250021 + i), Experiment_responders = ERs[a], ts = 3 + i)

```

```{r}
data <- sim_data[[1]]

```

#variance partitioning

```{r}
ar <- ASCA_decompose(d = apply(meta[,c(1,2)],2,as.factor), x = data, f = "growth_condition * time")

#unexpectedly different from PERMANOVA R2
norm(ar$residuals, type = "F")

# norm(ar$decomposition$Temperature, type = "F")

resids <- ar$residuals

# data <- resids
```

```{r}
spt <- pickSoftThreshold(data, moreNetworkConcepts = TRUE) 
# spt$powerEstimate
```

```{r}
softPower <- 6
adjacency <- adjacency(data, power = softPower)

TOM <- TOMsimilarity(adjacency)
```

```{r}
par(mfrow = c(1,2))
plot(spt$fitIndices[,1], -sign(spt$fitIndices[,3])*spt$fitIndices[,2],
      xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
      main = paste("Scale independence"));
  text(spt$fitIndices[,1], -sign(spt$fitIndices[,3])*spt$fitIndices[,2],
       labels=spt$fitIndices[,1],cex=0.9,col="red");
  # this line corresponds to using an R^2 cut-off of h
  abline(h=0.90,col="red")

# Mean connectivity as a function of the soft-thresholding power
plot(spt$fitIndices[,1], spt$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(spt$fitIndices[,1], spt$fitIndices[,5], labels=spt$fitIndices[,1], cex=0.9,col="red")

```

```{r}
TOM.dissimilarity <- 1-TOM
```

```{r}
geneTree <- hclust(as.dist(TOM.dissimilarity), method = "average")

sizeGrWindow(12,9)
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity",
labels = FALSE, hang = 0.04)
```

```{r}
Modules <- cutreeDynamic(dendro = geneTree, distM = TOM.dissimilarity, deepSplit = 2, pamRespectsDendro = FALSE, minClusterSize = 12)
table(Modules)

ModuleColors <- labels2colors(Modules) #assigns each module number a color
table(ModuleColors)

plotDendroAndColors(geneTree, ModuleColors,"Module",
dendroLabels = FALSE, hang = 0.03,
addGuide = TRUE, guideHang = 0.05,
main = "Gene dendrogram and module colors")
```

```{r}
MElist <- moduleEigengenes(data, colors = ModuleColors)
MEs <- MElist$eigengenes
head(MEs)
ME.dissimilarity = 1-cor(MElist$eigengenes, use="complete")
```

```{r}
METree = hclust(as.dist(ME.dissimilarity), method = "average") #Clustering eigengenes
par(mar = c(0,4,2,0)) #setting margin sizes
par(cex = 0.6);#scaling the graphic
plot(METree)
abline(h=.25, col = "red") #a height of .25 corresponds to correlation of .75

```

```{r}
merge <- mergeCloseModules(data, ModuleColors, cutHeight = .25)
# The merged module colors, assigning one color to each module
mergedColors = merge$colors
# Eigengenes of the new merged modules
mergedMEs = merge$newMEs

#check modules for POIs, URPs, 
#WGCNA also fails to discern between POIs and URPs under our simulated conditions as can be seen by module membership 
#although the miminum module number choice is arbitrarily chosen
tail(merge$colors,50)
```

```{r}

sKME <- signedKME(data, mergedMEs)

```

#selecting features

```{r}
# #calculate input data relationship to precalculated eigengenes
# sKME <- signedKME(data, mergedMEs)

#select coexpression candidates based on the cluster membership of a set of baits
#qbar target projection from MASCARA
q_bar <- colMeans(sKME[baits,])
R <- sKME[!rownames(sKME) %in% baits,]
R_TP <- (dot(t(R),q_bar)/dot(q_bar,q_bar))

#ordered list of candidates 
Candidates <- as.data.frame(R_TP[order(R_TP, decreasing = TRUE)])
```

<!-- ```{r} -->
<!-- plotDendroAndColors(geneTree, cbind(ModuleColors, mergedColors), -->
<!-- c("Original Module", "Merged Module"), -->
<!-- dendroLabels = FALSE, hang = 0.03, -->
<!-- addGuide = TRUE, guideHang = 0.05, -->
<!-- main = "Gene dendrogram and module colors for original and merged modules") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- nGenes = ncol(data) -->
<!-- nSamples = nrow(data) -->

<!-- #  -->
<!-- # module.trait.correlation = cor(mergedMEs, datTraits, use = "p") #p for pearson correlation coefficient -->
<!-- # module.trait.Pvalue = corPvalueStudent(module.trait.correlation, nSamples) #calculate the p-value associated with the correlation -->

<!-- ``` -->