### 2025-01-20 R/run_first/2.plot_pathway_net_example.Rmd

---
title: "network_plot"
author: "Fred White"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggrepel)
library(readxl)
library(extrafont)
library(cowplot)
library(graphlayouts)
library(gganimate)
library(animation)
library(gASCA)
library(matrixStats)
library(patchwork)

source("../DATA_SIM_FUNCS.R")
```

```{r}
# font_import() # download once 
#windowsFonts() # to check which are loaded

# renv::status()

```

```{r}
loadfonts(device = "win", quiet = TRUE)

###partial correlations

library(ppcor)
library(psych)

#
# plot residuals and full (with tweaks for problem demonstration)
baits <- paste0("X_",c(1985:1988))   #501:510
spikes <- paste0("X_",c(1989:2000))

baits <- c(baits,spikes[1:5])
spikes <- spikes[-c(1:5)]

meta <- cbind.data.frame(rep(c(1,-1), each = 12), rep(c(1:4), each = 3), rep(c(-1:-4,1:4), each = 3))
colnames(meta) <- c("growth_condition","time","interaction")
meta <- as.data.frame(apply(meta,2,factor))  #9

Experiment_responders <- 30

sim_data <- Create_Core_DEV_2(nreps = 3, meta = meta, 
                        EffectSize = c(X_a_ab = 5, time = 1, E = 1, mu = 1, Struc_E = 0.4),    # E = 10, Struc_E = 1
                        struc_resid = TRUE, e_sigma = c(1.5,0.8,0.0), a_sigma = c(1,0.7,0.6,0.1),
                        b_sigma = c(1,0.9), irr_spikes = TRUE, SCORE_SEED = 12, Experiment_responders = Experiment_responders,
                        struc_seed = 127, score_plot = FALSE, plot = FALSE)  #12

ar <- ASCA(sim_data[[1]], m = meta, Baits = baits, spikes = spikes, ncands = ncands, distance_calc = FALSE, Return_Model = FALSE)

residuals <- ar[[3]]
data <- sim_data[[1]]

cand_vals <- residuals[,1970:2000]    #rgs[which(rownames(rgs) %in% names(g_b)[1:50]),]
adj <- cor(cand_vals)

rownames(adj) <- colnames(cand_vals)
colnames(adj) <- colnames(cand_vals)

```

```{r}
set.seed(12)
pways <- sum(length(baits),length(spikes))
W <- matrix(sample(c(0,0,0,0,1),pways*pways, replace = T), nrow = pways)

adj[which(rownames(adj) %in% c(baits,spikes)),which(colnames(adj) %in% c(baits,spikes))] <- adj[which(rownames(adj) %in% c(baits,spikes)),which(colnames(adj) %in% c(baits,spikes))] * W

```

```{r}
set.seed(345)
W <- matrix(sample(c(1,0,0,0,0),15*15, replace = T), nrow = 15)
adj[-which(rownames(adj) %in% c(baits,spikes)),-which(colnames(adj) %in% c(baits,spikes))]  <- 0.9

adj[-which(rownames(adj) %in% c(baits,spikes)),-which(colnames(adj) %in% c(baits,spikes))] <- adj[-which(rownames(adj) %in% c(baits,spikes)),-which(colnames(adj) %in% c(baits,spikes))] * W

```

```{r}
#randomly sparsify p1 p2 connections

set.seed(22)

W <- matrix(sample(c(1,0,0,0,0),length(baits) * length(spikes), replace = T), nrow = length(spikes))

adj[which(rownames(adj) %in% spikes),which(colnames(adj) %in% baits)] <- adj[which(rownames(adj) %in% spikes),which(colnames(adj) %in% baits)] * W

```

```{r}
library(igraph)

g <- graph.adjacency(
  as.matrix(as.dist(adj)),
  mode="undirected",
  weighted=TRUE,
  diag=FALSE
)

g <- simplify(g, remove.multiple=TRUE, remove.loops=FALSE)

E(g)[which(E(g)$weight<0)]$association <- "-ve"
E(g)[which(E(g)$weight>0)]$association <- "+ve"

# Convert edge weights to absolute values
E(g)$weight <- abs(E(g)$weight)
```

```{r}
###
g <- delete_edges(g, E(g)[which(E(g)$weight<0.55)])

# Remove nodes with no edges
g <- delete.vertices(g, degree(g)==0)

V(g)$name <- V(g)$name

```

```{r}

coords <- layout_with_kk(g)  #to fix coordinates of nodes

```

```{r}

########
library(tidygraph)
library(ggraph)

Pathway  <- V(g)$name
```

```{r}
id <- c(1:length(Pathway ))

VIP <- rep(10,length(V(g)))

Pathway [which(Pathway  %in% baits)] <- "POI"
Pathway [which(Pathway  %in% spikes)] <- "Related"

Pathway [-which(Pathway  %in% c("POI","Related"))] <- "Unrelated"

temp <- paste0("X_",c(1991:1992,1989,1985))   #501:510

Feature <- names(V(g))
Feature[-which(Feature %in% temp)] <- "Other"

Feature[which(Pathway  %in% c("POI","Related"))] <- "Targets"

Feature[which(names(V(g)) %in% temp)] <- "Baits"

Feature <- factor(Feature, levels = c("Baits","Targets","Other"))

# Feature[which(names(V(g)) %in% spikes)] <- "Targets"

tidynet <- as_tbl_graph(g)

tidynet <- tidynet %>%
  activate(nodes) %>%
  mutate(VIP = VIP)

tidynet <- tidynet %>%
  activate(nodes) %>%
  mutate(Pathway  = Pathway )

tidynet <- tidynet %>%
  activate(nodes) %>%
  mutate(Feature = Feature)

tidynet <- tidynet %>%
  activate(nodes) %>%
  mutate(ID = names(V(g)))

colours = c("POI" = "purple", "Related" = "orange", "Unrelated" = "lightgreen")

# shape = c("Baits" = 18, "other" = 1)

edge_colours <- c("+ve" = "#091EB7", "-ve" = "blue")

realnet <- tidynet

realplot <- ggraph(tidynet, layout = coords) + 
  geom_edge_link(aes(colour = association), show.legend=FALSE) +  
  scale_edge_color_manual(values = edge_colours, guide = "none") +
  geom_node_point(
    aes(size = VIP, color = Pathway , shape = Feature), show.legend=FALSE        #CHANGE THIS
  ) +
  geom_node_point(
    aes(color = Pathway , shape = Feature)    #, shape = Baits
  ) +
  scale_color_manual(values = colours) + 
    scale_shape_manual(values = c(17,16,15)) +   #18,20

  theme_void()

realplot
```

```{r}
#only take nodes and layout from above

#remake graph or replace weights and edges with between group variance added

#remake adj with only names(V(g)) and calculate weights on data not residuals

#set specific weights in adj matrix   CHANGE THIS

cand_vals <- data[,which(colnames(data) %in% names(V(g)))]

adj <- cor(cand_vals)

rownames(adj) <- colnames(cand_vals)
colnames(adj) <- colnames(cand_vals)

adj[Pathway  == "Unrelated", Pathway  == "Unrelated"]  <- 1
adj[Pathway  %in% c("POI","Related"), Pathway  %in% c("POI","Related")] <- 1

adj[Pathway  %in% c("POI","Related"), Pathway  == "Unrelated"] <- -1
adj[Pathway  == "Unrelated", Pathway  %in% c("POI","Related")] <- -1
```

```{r}
#sparsify -1 connections

set.seed(22)

# adj[which(adj == -1)] <- adj[which(adj == -1)] * sample(c(1,0,0,0))

# adj[which(adj == -1, arr.ind = T)] <- adj[which(adj == -1,arr.ind = T)] * sample(c(1,0,0,0))

# W <- matrix(sample(c(1,1,0,0,0),length(baits) * length(spikes), replace = T), nrow = length(spikes))
# 
# 
# adj[which(rownames(adj) %in% spikes),which(colnames(adj) %in% baits)] <- adj[which(rownames(adj) %in% spikes),which(colnames(adj) %in% baits)] * W

```

```{r}
g <- graph.adjacency(
  as.matrix(as.dist(adj)),
  mode="undirected",
  weighted=TRUE,
  diag=FALSE
)

g <- simplify(g, remove.multiple=TRUE, remove.loops=FALSE)
```

```{r}

# Colour negative correlation edges as blue
E(g)[which(E(g)$weight<0)]$association <- "-ve"
# Colour positive correlation edges as red
E(g)[which(E(g)$weight>0)]$association <- "+ve"

# Convert edge weights to absolute values
E(g)$weight <- abs(E(g)$weight)

###
g <- delete_edges(g, E(g)[which(E(g)$weight<0.4)])

g <- delete.vertices(g, degree(g)==0)

V(g)$name <- V(g)$name

V(g)$shape <- "sphere"

#define correctly ordered type Pathway 
Pathway  <- V(g)$name

```

```{r}
id <- c(1:length(Pathway ))

VIP <- rep(10,length(V(g)))

Pathway [which(Pathway  %in% baits)] <- "POI"
Pathway [which(Pathway  %in% spikes)] <- "Related"

Pathway [-which(Pathway  %in% c("POI","Related"))] <- "Unrelated"

tidynet <- as_tbl_graph(g)

tidynet <- tidynet %>%
  activate(nodes) %>%
  mutate(VIP = VIP)

tidynet <- tidynet %>%
  activate(nodes) %>%
  mutate(Pathway  = Pathway )

colours = c("POI" = "purple", "Related" = "orange", "Unrelated" = "lightgreen")  
# shape = c("Baits" = 18, "other" = 1)
edge_colours <- c("+ve" = "black", "-ve" = "grey")

badnet <- tidynet

```

```{r}

badplot <- ggraph(tidynet, layout = coords) + 
  geom_edge_link(aes(colour = association, alpha = weight), show.legend=FALSE) +   #, linetype = "dashed"
  scale_edge_color_manual(values = edge_colours, guide = "none") +
  geom_node_point(
    aes(size = VIP, color = Pathway , shape = Feature), show.legend=FALSE
  ) +
  geom_node_point(
    aes(color = Pathway , shape = Feature)
  ) +
  scale_color_manual(values = colours) + 
      scale_shape_manual(values = c(17,16,15)) +   #18,20
  # scale_edge_linetype_manual(values = "dashed") +

  theme_void()

```

```{r}

tiff("../Fig1.tif", height = 3600, width = 4800, res = 700)
realplot + badplot + plot_annotation("Pathway Types", tag_levels = 'A') + plot_layout(guides = "collect")
dev.off()

```